Scalar Feedback based Joint Time-Frequency Precoder Interpolation for MIMO-OFDM Systems

### 2020-03-19 paper/main.tex
add changes, spelling correction

An underlying challenge in framing the predictive quantization and interpolation algorithms has been the high dimensional unitary structure of the SVD precoders. A large body of work utilizes the manifold geometry endowed by the unitary structure to frame such algorithms directly on the non-linear manifold ~\cite{schwarz2013adaptive,5946308,6891198,Gupt1905:Predictive,pitaval2013codebooks}. On the other hand, there are numerous works which utilize scalar parametrization of the unitary precoders, given by Givens rotations and Householder transformations \cite{4114278,4556174,lou2013comparison}. While the manifold based methods offer a clean solution and have been shown to be very effective in reducing the feedback overhead, they are analytically terse and
involve operations over higher dimensional manifolds. The scalar parametrization algorithms enable usage of the traditional algorithms to predict, interpolate and quantize over vector spaces, and hence are analytically simpler.
Such methods have also been implemented in 802.11 ac and ax standards
\cite{lou2013comparison,ieee80211}. We are not aware of performance comparisons
between these two bodies of work, thereby making it difficult to
quantify from past work whether performance gains/losses can accrue by operating over scalar parametrization instead of manifolds. In this paper, we propose a joint time-frequency interpolation approach that builds upon the existing parametric quantization techniques. 
One of the key results of our work is that the proposed scalar parametrization method actually outperforms an equivalent manifold based technique \cite{Gupt1905:Predictive}, at the same time being analytically more tractable.
With smaller quantization error as compared to manifold based approaches, the feedback
results in more accurate precoder tracking, along with simple 2D-interpolation over the
scalar parameters to optimize the precoders

Parametrization for precoder feedback have utilized adaptive delta
modulation effectively track precoding matrices both in the
time~\cite{4114278} and frequency domains~\cite{4556174}, albeit
separately. In our work, we perform prediction in time domain, and use
this for joint interpolation on the time frequency grid. Natural
wireless channels not only have temporal and frequency correlations,
but also possess joint time-frequency correlations.  Prior work
\cite{Gupt1905:Predictive} has framed an equivalent manifold based
approach to utilize joint time-frequency correlations for predictive
quantization and interpolation, and showed performance benefits
compared to utilizing time-frequency correlations independently
\cite{6891198,khaled2005quantized}.  Intuitively, these performance
benefits arise because utilization of joint correlations not only
exploits the existing independent correlations, but also the joint
correlations in time-frequency arising from the underlying channels.

Our key contributions are two-fold.  We first propose a method for exploiting the
joint time-frequency correlation of the precoder's scalar parameters
on the transmitter using the quantized feedback.
Secondly, instead of just comparing similar algorithms in the chosen body of work (viz. scalar parametrization), as has been the case with past work, we also compare our proposed algorithm with equivalent manifold based approaches, and show signficant performance benefits over manifold approaches.

@@

\subsection{Scalar Parametrization}

@@

Fig.~\ref{fig:error_decay} shows the rate at which error diminishes
with time for the proposed approach and direct quantization on the
Stiefel manifold as described in~\cite{Gupt1905:Predictive}. Since the
adaptation of independent scalar parameters causes the error to decay
rapidly, the error is lower and the convergence is faster, even though
the starting error is higher in the case of the scalar quantization
approach. The independence of the parameters enables fast
convergence. The Stiefel manifold based approach uses nested codebooks
that may be suboptimal when compared to the parametrization using
angles.

>>>>

An underlying challenge in framing the predictive quantization and interpolation algorithms has been the high dimensional unitary structure of the SVD precoders. A large body of work utilizes the manifold geometry endowed by the unitary structure to frame such algorithms directly on the non-linear manifold ~\cite{schwarz2013adaptive,5946308,6891198,Gupt1905:Predictive,pitaval2013codebooks}. On the other hand, there are numerous works which utilize scalar parameterization of the unitary precoders, given by Givens rotations and Householder transformations \cite{4114278,4556174,lou2013comparison}. While the manifold based methods offer a clean solution and have been shown to be very effective in reducing the feedback overhead, they are analytically terse and
involve operations over higher dimensional manifolds. The scalar parameterization algorithms enable usage of the traditional algorithms to predict, interpolate and quantize over vector spaces, and hence are analytically simpler.
Such methods have also been implemented in 802.11 ac and ax standards
\cite{lou2013comparison,ieee80211}. We are not aware of performance comparisons
between these two bodies of work, thereby making it difficult to
quantify from past work whether performance gains/losses can accrue by operating over scalar parameterization instead of manifolds. In this paper, we propose a joint time-frequency interpolation approach that builds upon the existing parametric quantization techniques. 
One of the key results of our work is that the proposed scalar parameterization method actually outperforms an equivalent manifold based technique \cite{Gupt1905:Predictive}, at the same time being analytically more tractable.
With smaller quantization error as compared to manifold based approaches, the feedback
results in more accurate precoder tracking, along with simple 2D-interpolation over the
scalar parameters to optimize the precoders

Past work in utilizing scalar parameterization for precoder feedback have utilized adaptive delta
modulation effectively track precoding matrices both in the
time~\cite{4114278} and frequency domains~\cite{4556174}, albeit
separately. In our work, we perform prediction in time domain, and use
this for joint interpolation on the time frequency grid. Natural
wireless channels not only have temporal and frequency correlations,
but also possess joint time-frequency correlations.  Prior work
\cite{Gupt1905:Predictive} has framed an equivalent manifold based
approach to utilize joint time-frequency correlations for predictive
quantization and interpolation, and showed performance benefits
compared to utilizing time-frequency correlations independently
\cite{6891198,khaled2005quantized}.  Intuitively, these performance
benefits arise because utilization of joint correlations not only
exploits the existing independent correlations, but also the joint
correlations in time-frequency arising from the underlying channels.

Our key contributions are two-fold.  We first propose a method for exploiting the
joint time-frequency correlation of the precoder's scalar parameters
on the transmitter using the quantized feedback.
Secondly, instead of just comparing similar algorithms in the chosen body of work (viz. scalar parameterization), as has been the case with past work, we also compare our proposed algorithm with equivalent manifold based approaches, and show signficant performance benefits over manifold approaches.

@@

\subsection{Scalar Parameterization}

@@

Fig.~\ref{fig:error_decay} shows the rate at which error diminishes
with time for the proposed approach and direct quantization on the
Stiefel manifold as described in~\cite{Gupt1905:Predictive}. Since the
adaptation of independent scalar parameters causes the error to decay
rapidly, the error is lower and the convergence is faster, even though
the starting error is higher in the case of the scalar quantization
approach. The independence of the parameters enables fast
convergence. The Stiefel manifold based approach uses nested codebooks
that may be suboptimal when compared to the parameterization using
angles.
